Morphological Operations Applied to Digital Art Restoration

### 2014-04-10 LaTeX Paper/paper.tex
Minor additions/corrections/adjustments to paper...

\title{Morphological Operations Applied to Crack Detection in Digital Art Restoration}

@@

This paper breaks the digital art restoration process into three steps: edge detection, crack detection, and crack removal. Section~\ref{edge detection} covers edge detection by examining an optimal edge detector known as the Canny edge detection algorithm. Section~\ref{morphological operations} provides an examination of the four basic morphological operators of erosion, dilation, opening, and closing. This information is necessary to understanding Section~\ref{crack detection}, which covers crack detection, and is the main focus of this paper. Section~\ref{inpainting} wraps up the digital art restoration process in covering the crack removal process (known as \textit{inpainting}). The remaining sections provide a summary. Section~\ref{results} reviews the results obtained by various methods of digital art restoration, again with a specific focus on the success of various morphological operators in the crack detection stage, and Section~\ref{conclusion} provides a summary of the information provided in this paper, as well as discusses possible further areas of interest and study.

@@

The Canny algorithm is based upon three criteria, which are as follows~\cite{Canny:1986, CannySummarized}:
\begin{enumerate}
\item Accuracy - The edge detection algorithm should have a low error rate. This implies that there should be neither false positives (edges detected where they do not exist) nor false negatives (edges not detected where they do exist).
\item Localization - The edge detection algorithm should minimize the distance between the edge detected and the actual edge.
\item Uniqueness - The edge detection algorithm should have only one response to a single edge.
\end{enumerate}

In order to fulfill these criteria, the Canny algorithm first uses a Gaussian filter to smooth the image. The Gaussian filter functions in a manner somewhat similar to the structuring element used in mathematical morphology, discussed in Section~\ref{morphological operations}. Essentially, the filter passes over each pixel in turn, blending each into the surrounding pixels. The complete result is a slightly blurred version of the original image with no isolated, ``noisy'' pixels (pixels that exhibit a significant jump in intensity from all surrounding pixels).

@@

\section{Morphological Operations}\label{morphological operations}
Following optional edge detection is crack detection, the focus of this paper. However, in order to understand this stage, it is first necessary to describe the basic principles of mathematical morphology. Typically, morphology is used in processing binary (black and white) images, although there are also variations used for greyscale images. This paper will examine both variations. Morphological functions take two inputs. The first input is the image to be processed. In binary morphology, the input image is divided into foreground (typically white) and background (typically black) regions. In greyscale morphology, the image is treated as a three-dimensional object, where higher areas are brighter (white) and lower areas are darker (black). The second input is a structuring element, a (typically small, in comparison to the image) set of coordinate points. An example of a 3x3 square structuring element with the origin located at the center in comparison to both a 20x30 binary image and a 32x11 greyscale image is presented in Figure~\ref{structuring element figure}. The structuring element is then used to modify the input image. The image that results from the morphological operation is determined by the shape, size, and point of origin of the structuring element~\cite{MorphologyWikiAnonymous, MorphologyBook:2000, MorphologyPaper:1987, MorphologyWiki}. Subsections~\ref{erosion}, \ref{dilation}, \ref{opening}, and \ref{closing} explain the fundamental operations of mathematical morphology.
\begin{figure}
\centering
\includegraphics[width=3in,trim={0 6.5in 0 0},clip]{structuring_element}
\caption{Structuring Elements: Top: Structuring Element. Bottom Left: Structuring Element in Comparison to Binary Image. Bottom Right: Structuring Element in Comparison to Greyscale Image (Above) and Cross-Section (Below).}
\label{structuring element figure}
\end{figure}

@@

Greyscale erosion follows the same basic pattern, but is somewhat more complex. Recall that in greyscale morphology, the image is viewed as a three-dimensional object, with ``hills'' corresponding to higher intensities and ``valleys'' indicating lower intensities. Here, the origin of the structuring element is placed in line with a pixel within the object. If every point within the structuring element is contained within the object, the pixel lined up with the origin is left unchanged. If at least one point within the structuring element is outside the object, the pixel lined up with the origin of the structuring element is removed. This process is repeated for every pixel within the object~\cite{MorphologyPaper:1987, MorphologyWiki}. An example of greyscale erosion performed on a cross-section of a 32x11 image using a 3x3 square structuring element with the origin located at the center is presented in Figure~\ref{greyscale erosion figure}.
\begin{figure}
\centering
\includegraphics[width=3in,trim={0 4in 0 0},clip]{greyscale_erosion}
\caption{Greyscale Erosion: Top: Original Image. Center Left: Cross-Section of Image Surface. Center Center: Erosion Marked. Center Right: Erosion of Surface. Bottom: Results of Erosion.}
\label{greyscale erosion figure}
\end{figure}

@@

Like greyscale erosion, greyscale dilation is a more complex variation on its binary counterpart. Once again, the origin of the structuring element is placed in line with a pixel within the three-dimensional object that represents the original image. If every point within the structuring element is contained within the object, the object is left unchanged. If at least one point within the structuring element is outside the object, the regions outside of the object lined up with points within the structuring element are set to be a part of the object. This process is repeated for every pixel within the object~\cite{MorphologyPaper:1987, MorphologyWiki}. An example of greyscale dilation performed on a cross-section of a 32x11 image using a 3x3 square structuring element with the origin located at the center is presented in Figure~\ref{greyscale dilation figure}.
\begin{figure}
\centering
\includegraphics[width=3in,trim={0 4in 0 0},clip]{greyscale_dilation}
\caption{Greyscale Dilation: Top: Original Image. Center Left: Cross-Section of Image Surface. Center Center: Dilation Marked. Center Right: Dilation of Surface. Bottom: Results of Dilation.}
\label{greyscale dilation figure}
\end{figure}

@@

\subsection{Top-Hat Transform}
The morphological filter most frequently used for detecting cracks within digitalized works of art is the top-hat transform. There are two variations on this filter, the black top-hat transform and white top-hat transform, which will be discussed in Subsubsections~\ref{black top-hat} and \ref{white top-hat} respectively.

\subsubsection{Black Top-Hat}\label{black top-hat}
The black top-hat transform (also referred to as the closing top-hat transform) is used for detecting darker details on a lighter background within greyscale images~\cite{Altarpiece:2013, TopHat:2010}. This transform is defined as the difference between the closing of the original image by a specified structuring element and the original image itself, and is denoted by the equation
\begin{equation*}
BTH = (f \bullet s) - f
\end{equation*}
where \textit{f} is the original image and \textit{s} is the structuring element. This transformation produces a greyscale image with the desired details enhanced.

\subsubsection{White Top-Hat}\label{white top-hat}
The white top-hat transform (also referred to as the opening top-hat transform or, sometimes, the bottom-hat transform), is used for detecting lighter details on a darker background within greyscale images~\cite{Altarpiece:2013, TopHat:2010}. This transform is defined as the difference between the original image itself and the opening of the original image by a specified structuring element, and is denoted by the equation
\begin{equation*}
WTH = f - (f \circ s)
\end{equation*}
where \textit{f} is the original image and \textit{s} is the structuring element. Like its counterpart, this transformation produces a greyscale image with the desired details enhanced.

\subsubsection{Multiscale Top-Hat}\label{multiscale top-hat}
Subsubsections~\ref{black top-hat} and \ref{white top-hat} explain classical top-hat transforms. While these are popular and frequently used, they are not perfect. Paintings often contain details (such as fine brush strokes) which are misinterpreted as cracks by these filters. The multiscale morphological top-hat transform may be used to counteract this, as well as to detect cracks of varying sizes~\cite{Altarpiece:2013}. This transformation is executed by performing a series of classical top-hat transforms with structuring elements of various shapes and sizes (Cornelis et al use square structuring elements ``ranging from 3x3 to \textit{n}x\textit{n} pixels, where \textit{n} depends on the width of the crack to be detected''~\cite{Altarpiece:2013}). The resulting images, or crack maps, are then further processed to remove isolated groups of pixels and bridge pixel gaps.

@@

The statistics reported are summarized in Table~\ref{results table}. While the true positive rate, or recall, recorded by Karianakis and Maragos is proportional to that of the top-hat transform when the percentage of edge information lost is low, the alternative very quickly losses any comparability as the percentage of edge information lost increases. And while the precision does increase as edge information is lost, even the highest precision demonstrated by the alternative does not compare with the precision of the top-hat transform. It is therefore reasonable to conclude that the top-hat transform, in both its classical and multiscale forms, significantly outperforms the alternative implemented by Karianakis and Maragos.
\begin{table*}
\centering
%\resizebox{8.5cm}{!}{
\begin{tabular}{|l|l|r|r|r|r|r|r|r|}
\hline
\textbf{Method} & \textbf{Classification} & \textbf{\textit{tp}} & \textbf{\textit{fn}} & \textbf{\textit{tn}} & \textbf{\textit{fp}} & \textbf{\textit{TP} (or \textit{R})} & \textbf{\textit{FP}} & \textbf{\textit{P}}\\ \hline
\multirow{8}{*}{Top-Hat Transform}
& Crack Thickness - Thin & 220 & 30 & 230 & 20 & 0.880 & 0.080 & 0.917\\ \cline{2-9}
& Crack Thickness - Medium & 232 & 18 & 231 & 19 & 0.928 & 0.076 & 0.924\\ \cline{2-9}
& Crack Thickness - Thick & 235 & 15 & 238 & 12 & 0.940 & 0.048 & 0.951\\ \cline{2-9}
& Number of Cracks - Few & 242 & 8 & 245 & 5 & 0.968 & \textbf{0.020} & 0.980\\ \cline{2-9}
& Number of Cracks - Medium & 245 & 5 & 241 & 9 & \textbf{0.980} & 0.036 & 0.965\\ \cline{2-9}
& Number of Cracks - Many & 243 & 7 & 243 & 7 & 0.972 & 0.028 & 0.972\\ \cline{2-9}
& Crack Connectivity - Low & 215 & 35 & 219 & 31 & \textbf{0.860} & \textbf{0.124} & 0.874\\ \cline{2-9}
& Crack Connectivity - High & 218 & 32 & 221 & 29 & 0.872 & 0.116 & 0.883\\ \hline
\multirow{3}{*}{Alternative Method}
& Edge Information Lost - 1\% & - & - & - & - & \textbf{0.932} & - & \textbf{0.497}\\ \cline{2-9}
& Edge Information Lost - 30\% & - & - & - & - & 0.857 & - & 0.594\\ \cline{2-9}
& Edge Information Lost - 70\% & - & - & - & - & \textbf{0.530} & - & \textbf{0.704}\\ \hline
\end{tabular}
%}
\caption{Crack Detection Results by Methodology: Data Specified in Text Emphasized, Modified from~\cite{Statistics:2013, Theran:2013}.}
\label{results table}
\end{table*}

@@

\section{Acknowledgments}
Thanks to Peter Dolan for his input, help, and advice, and to Elena Machkasova for her recommendations.

>>>>

\title{Morphological Operations Applied to Digital Art Restoration}

@@

This paper breaks the digital art restoration process into three steps: edge detection, crack detection, and crack removal. Section~\ref{edge detection} covers edge detection by examining an optimal edge detector known as the Canny edge detection algorithm. Section~\ref{morphological operations} provides an examination of the four basic morphological operators of erosion, dilation, opening, and closing. This information is necessary for understanding Section~\ref{crack detection}, which covers crack detection, and is the main focus of this paper. Section~\ref{inpainting} wraps up the digital art restoration process in covering the crack removal process (known as \textit{inpainting}). The remaining sections provide a summary. Section~\ref{results} reviews the results obtained by various methods of digital art restoration, again with a specific focus on the success of various morphological operators in the crack detection stage, and Section~\ref{conclusion} provides a summary of the information provided in this paper, as well as discusses possible further areas of interest and study.

@@

The Canny algorithm is based upon three criteria, which are as follows~\cite{Canny:1986, CannySummarized}:
\begin{enumerate}
\item Accuracy - The edge detection algorithm should have a low error rate. This implies that there should be neither false positives (edges detected where they do not exist) nor false negatives (edges not detected where they do exist).
\item Localization - The edge detection algorithm should minimize the distance between the edge detected and the actual edge.
\item Uniqueness - The edge detection algorithm should have only one response to a single edge. That is, one edge, or jump in intensity, should equal one response.
\end{enumerate}

In order to fulfill these criteria, the Canny algorithm first uses a Gaussian filter to smooth the image. This is achieved by passing the filter over each pixel in turn, blending each into the surrounding pixels. The complete result is a slightly blurred version of the original image with no isolated, ``noisy'' pixels (pixels that exhibit a significant jump in intensity from all surrounding pixels).

@@

\section{Morphological Operations}\label{morphological operations}
Following optional edge detection is crack detection, the focus of this paper. However, in order to understand this stage, it is first necessary to describe the basic principles of mathematical morphology. Typically, morphology is used in processing binary (black and white) images, although there are also variations used for greyscale images. This paper will examine both variations. Morphological functions take two inputs. The first input is the image to be processed. In binary morphology, the input image is divided into foreground (typically white) and background (typically black) regions. In greyscale morphology, the image is treated as a three-dimensional object, where higher areas are brighter (white) and lower areas are darker (black). The second input is a structuring element, a (typically small, in comparison to the image) set of coordinate points. An example of a 3x3 square structuring element with the origin located at the center in comparison to a 20x30 binary image is presented in Figure~\ref{structuring element figure}. The structuring element is then used to modify the input image. The image that results from the morphological operation is determined by the shape, size, and point of origin of the structuring element~\cite{MorphologyWikiAnonymous, MorphologyBook:2000, MorphologyPaper:1987, MorphologyWiki}. Subsections~\ref{erosion}, \ref{dilation}, \ref{opening}, and \ref{closing} explain the fundamental operations of mathematical morphology.
\begin{figure}
\centering
\includegraphics[width=1.75in,trim={0 6.5in 4in 0},clip]{structuring_element_placement}
\caption{Structuring Element (Letter Marks Origin): A: Contained in Foreground Region. B: Intersecting Foreground and Background Regions, Origin in Foreground Region. C: Contained in Background Region. D: Intersecting Foreground and Background Regions, Origin in Background Region.}
\label{structuring element figure}
\end{figure}

@@

Greyscale erosion follows the same basic pattern, but is somewhat more complex. Recall that in greyscale morphology, the image is viewed as a three-dimensional object, with ``hills'' corresponding to higher intensities and ``valleys'' indicating lower intensities. Here, the origin of the structuring element is aligned with a pixel within the object. If every point within the structuring element is contained within the object, the pixel aligned with the origin is left unchanged. If at least one point within the structuring element is outside the object, the pixel aligned with the origin of the structuring element is removed. This process is repeated for every pixel within the object~\cite{MorphologyPaper:1987, MorphologyWiki}. An example of greyscale erosion performed on a cross-section of a 32x11 image using a 3x3 square structuring element with the origin located at the center is presented in Figure~\ref{greyscale erosion figure}.
\begin{figure}
\centering
\includegraphics[width=3in,trim={0 4in 0 0},clip]{greyscale_erosion}
\caption{Greyscale Erosion: Top: Original Image. Center Left: Cross-Section of Image Surface. Center Center: Erosion Marked. Center Right: Erosion of Surface. Bottom: Results of Erosion.}
\label{greyscale erosion figure}
\end{figure}

@@

Like greyscale erosion, greyscale dilation is a more complex variation on its binary counterpart. Once again, the origin of the structuring element is aligned with a pixel within the three-dimensional object that represents the original image. If every point within the structuring element is contained within the object, the object is left unchanged. If at least one point within the structuring element is outside the object, the regions outside of the object aligned with points within the structuring element are set to be a part of the object. This process is repeated for every pixel within the object~\cite{MorphologyPaper:1987, MorphologyWiki}. An example of greyscale dilation performed on a cross-section of a 32x11 image using a 3x3 square structuring element with the origin located at the center is presented in Figure~\ref{greyscale dilation figure}.
\begin{figure}
\centering
\includegraphics[width=3in,trim={0 4in 0 0},clip]{greyscale_dilation}
\caption{Greyscale Dilation: Top: Original Image. Center Left: Cross-Section of Image Surface. Center Center: Dilation Marked. Center Right: Dilation of Surface. Bottom: Results of Dilation.}
\label{greyscale dilation figure}
\end{figure}

@@

\subsection{Top-Hat Transform}
The morphological filter most frequently used for detecting cracks within digitalized works of art is the top-hat transform. There are two variations on this filter, the black top-hat transform and white top-hat transform, which will be discussed in Subsections~\ref{black top-hat} and \ref{white top-hat} respectively.

\subsubsection{Black Top-Hat}\label{black top-hat}
The black top-hat transform (also referred to as the closing top-hat transform) is used for detecting darker details on a lighter background within greyscale images~\cite{Altarpiece:2013, TopHat:2010}. This transform is defined as the difference in pixel intensity between the closing of the original image by a specified structuring element and the original image itself, and is denoted by the equation
\begin{equation*}
BTH = (f \bullet s) - f
\end{equation*}
where \textit{f} is the original image and \textit{s} is the structuring element. This transformation produces a greyscale image with the desired details enhanced.

\subsubsection{White Top-Hat}\label{white top-hat}
The white top-hat transform (also referred to as the opening top-hat transform or, sometimes, the bottom-hat transform), is used for detecting lighter details on a darker background within greyscale images~\cite{Altarpiece:2013, TopHat:2010}. This transform is defined as the difference in pixel intensity between the original image itself and the opening of the original image by a specified structuring element, and is denoted by the equation
\begin{equation*}
WTH = f - (f \circ s)
\end{equation*}
where \textit{f} is the original image and \textit{s} is the structuring element. Like its counterpart, this transformation produces a greyscale image with the desired details enhanced.

\subsubsection{Multiscale Top-Hat}\label{multiscale top-hat}
Subsections~\ref{black top-hat} and \ref{white top-hat} explain classical top-hat transforms. While these are popular and frequently used, they are not perfect. Paintings often contain details (such as fine brush strokes) which are misinterpreted as cracks by these filters. The multiscale morphological top-hat transform may be used to counteract this, as well as to detect cracks of varying sizes~\cite{Altarpiece:2013}. This transformation is executed by performing a series of classical top-hat transforms with structuring elements of various shapes and sizes (Cornelis et al use square structuring elements ``ranging from 3x3 to \textit{n}x\textit{n} pixels, where \textit{n} depends on the width of the crack to be detected''~\cite{Altarpiece:2013}). The resulting images, or crack maps, are then further processed to remove isolated groups of pixels and bridge pixel gaps.

@@

The statistics reported are summarized in Table~\ref{results table}. While the true positive rate, or recall, recorded by Karianakis and Maragos is proportional to that of the top-hat transform when the percentage of edge information lost is low, the alternative very quickly losses any comparability as the percentage of edge information lost increases. And while the precision does increase as edge information is lost, even the highest precision demonstrated by the alternative does not compare with the precision of the top-hat transform. It is therefore reasonable to conclude that the top-hat transform, in both its classical and multiscale forms, significantly outperforms the alternative implemented by Karianakis and Maragos.
\begin{table*}
\centering
%\resizebox{8.5cm}{!}{
\begin{tabular}{|l|l|r|r|r|r|r|r|r|}
\hline
\textbf{Method} & \textbf{Classification} & \textbf{\textit{tp}} & \textbf{\textit{fn}} & \textbf{\textit{tn}} & \textbf{\textit{fp}} & \textbf{\textit{TP} (or \textit{R})} & \textbf{\textit{FP}} & \textbf{\textit{P}}\\ \hline
\multirow{8}{*}{Top-Hat Transform}
& Crack Thickness - Thin & 220 & 30 & 230 & 20 & 0.880 & 0.080 & 0.917\\ \cline{2-9}
& Crack Thickness - Medium & 232 & 18 & 231 & 19 & 0.928 & 0.076 & 0.924\\ \cline{2-9}
& Crack Thickness - Thick & 235 & 15 & 238 & 12 & 0.940 & 0.048 & 0.951\\ \cline{2-9}
& Number of Cracks - Few & 242 & 8 & 245 & 5 & 0.968 & \textbf{0.020} & 0.980\\ \cline{2-9}
& Number of Cracks - Medium & 245 & 5 & 241 & 9 & \textbf{0.980} & 0.036 & 0.965\\ \cline{2-9}
& Number of Cracks - Many & 243 & 7 & 243 & 7 & 0.972 & 0.028 & 0.972\\ \cline{2-9}
& Crack Connectivity - Low & 215 & 35 & 219 & 31 & \textbf{0.860} & \textbf{0.124} & 0.874\\ \cline{2-9}
& Crack Connectivity - High & 218 & 32 & 221 & 29 & 0.872 & 0.116 & 0.883\\ \hline
\multirow{3}{*}{Alternative Method}
& Edge Information Lost - 1\% & - & - & - & - & \textbf{0.932} & - & \textbf{0.497}\\ \cline{2-9}
& Edge Information Lost - 30\% & - & - & - & - & 0.857 & - & 0.594\\ \cline{2-9}
& Edge Information Lost - 70\% & - & - & - & - & \textbf{0.530} & - & \textbf{0.704}\\ \hline
\end{tabular}
%}
\caption{Crack Detection Results by Methodology: Data Specified in Text Emphasized, Modified from~\cite{Statistics:2013, Theran:2013}. \textit{tp} and \textit{fn} Total 500, Same for \textit{tn} and \textit{fp}.}
\label{results table}
\end{table*}

@@

\section{Acknowledgments}
Thanks to Peter Dolan for his input, help, and advice, to Elena Machkasova for her corrections and improvements, and to Brian Goslinga for his recommendations.
